### 2018-01-09 master.tex
Kolya's changes (Jan 9)
Auger figure, author list

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% This is a (brief) model paper using the achemso class
%% The document class accepts keyval options, which should include
%% the target journal and optionally the manuscript type.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[journal=jacsat,manuscript=article]{achemso}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional packages needed here.  Only include packages
%% which are essential, to avoid problems later. Do NOT use any
%% packages which require e-TeX (for example etoolbox): the e-TeX
%% extensions are not currently available on the ACS conversion
%% servers.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
\usepackage[T1]{fontenc}       % Use modern font encodings

\usepackage{color}
\usepackage{xr}
\externaldocument{suppinfo}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% If issues arise when submitting your manuscript, you may want to
%% un-comment the next line.  This provides information on the
%% version of every file you have used.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\listfiles

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional macros here.  Please use \newcommand* where
%% possible, and avoid layout-changing macros (which are not used
%% when typesetting).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand*\mycommand[1]{\texttt{\emph{#1}}}

\newcommand{\cli}{Cl$^{-}$}
\newcommand{\ki}{K$^{+}$}

\newcommand{\tsveta}[1]{\textcolor{red}{#1}}
\newcommand{\nico}[1]{\textcolor{red}{#1}}
\newcommand{\kolya}[1]{\textcolor{blue}{#1}}
\newcommand{\denis}[1]{\textcolor{green}{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Meta-data block
%% ---------------
%% Each author should be given as a separate \author command.
%%
%% Corresponding authors should have an e-mail given after the author
%% name as an \email command. Phone and fax numbers can be given
%% using \phone and \fax, respectively; this information is optional.
%%
%% The affiliation of authors is given after the authors; each
%% \affiliation command applies to all preceding authors not already
%% assigned an affiliation.
%%
%% The affiliation takes an option argument for the short name.  This
%% will typically be something like "University of Somewhere".
%%
%% The \altaffiliation macro should be used for new address, etc.
%% On the other hand, \alsoaffiliation is used on a per author basis
%% when authors are associated with multiple institutions.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Tsveta Miteva}
\affiliation{Sorbonne Universit\'{e}s, UPMC Univ Paris 06, UMR 7614, Laboratoire de Chimie Physique Mati\`{e}re et Rayonnement, F-75005 Paris, France}

\author{Nikolai Kryzhevoi}
\affiliation{Theoretische Chemie, Physikalisch-Chemisches Institut, Universit\"at Heidelberg, Im Neuenheimer Feld 229, D-69120 Heidelberg, Germany}

\author{Nicolas Sisourat}
\affiliation{Sorbonne Universit\'{e}s, UPMC Univ Paris 06, UMR 7614, Laboratoire de Chimie Physique Mati\`{e}re et Rayonnement, F-75005 Paris, France}

\author{Ch. Nicolas}
\affiliation{Sorbonne Universit\'{e}s, UPMC Univ Paris 06, UMR 7614, Laboratoire de Chimie Physique Mati\`{e}re et Rayonnement, F-75005 Paris, France}

\author{Wandared Pokapanich}
\affiliation{Sorbonne Universit\'{e}s, UPMC Univ Paris 06, UMR 7614, Laboratoire de Chimie Physique Mati\`{e}re et Rayonnement, F-75005 Paris, France}

\author{Th. Saisopa}
\affiliation{Sorbonne Universit\'{e}s, UPMC Univ Paris 06, UMR 7614, Laboratoire de Chimie Physique Mati\`{e}re et Rayonnement, F-75005 Paris, France}

\author{P. Songsiriritthigul}
\affiliation{Sorbonne Universit\'{e}s, UPMC Univ Paris 06, UMR 7614, Laboratoire de Chimie Physique Mati\`{e}re et Rayonnement, F-75005 Paris, France}

\author{Y. Rattanachai}
\affiliation{Sorbonne Universit\'{e}s, UPMC Univ Paris 06, UMR 7614, Laboratoire de Chimie Physique Mati\`{e}re et Rayonnement, F-75005 Paris, France}

\author{Andreas Dreuw}
\affiliation{Interdisciplinary Center for Scientific Computing, Ruprecht-Karls University, Im Neuenheimer Feld 205A, D-69120 Heidelberg, Germany}

\author{Jan Wenzel}
\affiliation{Interdisciplinary Center for Scientific Computing, Ruprecht-Karls University, Im Neuenheimer Feld 205A, D-69120 Heidelberg, Germany}

\author{Ralph P\"{u}ttner}
\affiliation{Sorbonne Universit\'{e}s, UPMC Univ Paris 06, UMR 7614, Laboratoire de Chimie Physique Mati\`{e}re et Rayonnement, F-75005 Paris, France}

\author{J\'{e}r\^ome Palaudoux}
\affiliation{Sorbonne Universit\'{e}s, UPMC Univ Paris 06, UMR 7614, Laboratoire de Chimie Physique Mati\`{e}re et Rayonnement, F-75005 Paris, France}

\author{Gunnar \"{O}hrwall}
\affiliation{Sorbonne Universit\'{e}s, UPMC Univ Paris 06, UMR 7614, Laboratoire de Chimie Physique Mati\`{e}re et Rayonnement, F-75005 Paris, France}

\author{Lorenz S. Cederbaum}
\affiliation{Theoretische Chemie, Physikalisch-Chemisches Institut, Universit\"at Heidelberg, Im Neuenheimer Feld 229, D-69120 Heidelberg, Germany}

\author{Jean Pascal Rueff}
\affiliation{Sorbonne Universit\'{e}s, UPMC Univ Paris 06, UMR 7614, Laboratoire de Chimie Physique Mati\`{e}re et Rayonnement, F-75005 Paris, France}
\affiliation{Synchrotron SOLEIL, l`Orme des Merisiers, Saint-Aubin, F-91192 Gif-sur-Yvette Cedex, France}

\author{Denis C\'{e}olin}
\email{denis.ceolin@synchrotron-soleil.fr}
\affiliation{Synchrotron SOLEIL, l`Orme des Merisiers, Saint-Aubin, F-91192 Gif-sur-Yvette Cedex, France}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[]
  {The electronic structure of aqueous KCl revealed by X-ray absorption and Auger electron spectroscopy
\\  
  The all-seeing eye of Auger electron spectroscopy: a study on aqueous KCl}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Some journals require a list of abbreviations or keywords to be
%% supplied. These should be set up here, and will be printed after
%% the title and author information, if needed.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\abbreviations{AES,XAS}
\keywords{Solvated ions, Auger spectroscopy, X-ray absorption spectroscopy}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The manuscript does not need to include \maketitle, which is
%% executed automatically.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "tocentry" environment can be used to create an entry for the
%% graphical table of contents. It is given here as some journals
%% require that it is printed as part of the abstract page. It will
%% be automatically moved as appropriate.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{tocentry}

\includegraphics[scale=0.82]{figures/graphical.png}
%Some journals require a graphical entry for the Table of Contents.
%This should be laid out ``print ready'' so that the sizing of the
%text is correct.
%
%Inside the \texttt{tocentry} environment, the font used is Helvetica
%8\,pt, as required by \emph{Journal of the American Chemical
%Society}.
%
%The surrounding frame is 9\,cm by 3.5\,cm, which is the maximum
%permitted for  \emph{Journal of the American Chemical Society}
%graphical table of content entries. The box will not resize if the
%content is too big: instead it will overflow the edge of the box.
%
%This box and the associated title will always be printed on a
%separate page at the end of the document.

\end{tocentry}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The abstract environment will automatically gobble the contents
%% if an abstract is not used by the target journal.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
 \input{abstract.tex}
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\input{introduction.tex}

\input{methods.tex}

\input{results.tex}

\input{conclusion.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "Acknowledgement" section can be given in all manuscript
%% classes.  This should be given within the "acknowledgement"
%% environment, which will make the correct section or running title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Acknowledgement}

\begin{acknowledgement}

\tsveta{We thank Prof.\ Nobuhiro Kosugi and Dr.\ Matja\v{z} \v{Z}itnik for the fruitful discussions?.}
Experiments were performed at the GALAXIES beamline, SOLEIL Synchrotron, France (Proposal No. 20140160). The authors are grateful to the SOLEIL staff for assistance during the beamtime. This project has received funding from the Research Executive Agency (REA) under the European Union$'$s Horizon 2020 research and innovation programme Grant agreement No 705515. Campus France and the PHC SIAM exchange program are acknowledged for financial support.

\end{acknowledgement}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The same is true for Supporting Information, which should use the
%% suppinfo environment.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{suppinfo}

\begin{itemize}
  \item suppinfo.pdf: contains the radial density distributions of the core excited states of the bare ions.
\end{itemize}

\end{suppinfo}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The appropriate \bibliography command should be placed here.
%% Notice that the class file automatically sets \bibliographystyle
%% and also names the section correctly.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{Bibliography}

\end{document}