### 2018-02-19 conclusion.tex
feb 16 (ralph's corrections included)
Ralph's corrections, theoretical figures changed, 1s -> nl replaced by
1s-1 nl

\section{Conclusion}\label{sec:concl}

Using a combination of x-ray absorption and Auger electron spectroscopy in the tender x-ray regime, we studied the electronic structure of aqueous solution of KCl at the K-edges of both K and Cl. The Auger electron spectra of both ions as a function of photon energy exhibit features of normal as well as resonant Auger processes. To interpret the resonant Auger features in the experimental spectrum, we performed {\it ab initio} calculations on microsolvated clusters of \ki~and \cli. Our calculations show that the energy ordering of the 3d and 4p virtual orbitals of \cli~is inverted compared to \ki, and also that the energy splitting between the bright 1s$^{-1}$4p and dark 1s$^{-1}$3d core excited states is larger in the chloride case. Thus, the energetic proximity of the 3d and 4p orbitals in the bare \ki~ion results in the dipole forbidden 1s$^{-1}$3d state acquiring intensity in a solution as a result of mixing with the dipole allowed 1s$^{-1}$4p excitation. The spectator Auger decay of this state produces an additional dispersive feature which is manifested as a separate peak in the Auger electron spectrum at high kinetic energies. In the case of \cli~the 1s$^{-1}$4p and 1s$^{-1}$3d core excited states do not interact, and therefore, only fingerprints of the population and Auger decay of the dipole allowed 1s$^{-1}$4p excitation are observed in the spectrum. Moreover, using the core-hole clock method we estimated the time of delocalization of the core excited electron at the pre-edge region of \cli$_{\text{aq}}$. Our results show that in this case, the resonant Auger decay and the delocalization of the excited electron occur on a comparable timescale. In the case of \ki$_{\text{aq}}$, we cannot make an accurate estimate, however, one can expect a much less efficient delocalization of the core excited electron.

Our work shows that the combination of x-ray absorption and resonant Auger spectroscopy is a sensitive probe of the electronic structure of solvated ions. The reported results are an important first step in the study of the electronic decay processes following photoabsorption in the tender x-ray regime. The final Auger states inevitably undergo subsequent multiple electronic decay processes, such as ICD and ETMD. As a result of these processes, the surrounding environment becomes damaged, i.e.\ it is ionized, and moreover, slow electrons are emitted, which can inflict further damage upon the surroundings. The magnitude of the damage and the energies of the emitted electrons depend on the initial Auger step, and can therefore be controlled by varying the photon energy. Consequently, the results of this work can have implications in revealing the mechanisms of radiation damage in biologically relevant systems.