### 2018-04-23 conclusion.tex
apr 23
preparation of jpcl draft started

%\section{Conclusion}\label{sec:concl}

In summary, we studied the electronic structure of aqueous solution of KCl at the K-edges of both K and Cl using a combination of x-ray absorption and Auger electron spectroscopy in the tender x-ray regime, and {\it ab initio} calculations. The Auger electron spectra of both ions as a function of photon energy exhibit features of normal as well as resonant Auger processes. The latter process proceeds differently for aqueous K$^{+}$ and Cl$^{-}$ due to the population of the dipole forbidden K$^{+}$ 1s$^{-1}$3d state in a solution. The spectator Auger decay of this state produces an additional dispersive feature which is manifested as a separate peak in the Auger electron spectrum. In the case of \cli~only fingerprints of the population and Auger decay of the dipole allowed 1s$^{-1}$4p excitation are observed in the spectrum.

%Using a combination of x-ray absorption and Auger electron spectroscopy in the tender x-ray regime, we studied the electronic structure of aqueous solution of KCl at the K-edges of both K and Cl. The Auger electron spectra of both ions as a function of photon energy exhibit features of normal as well as resonant Auger processes. To interpret the resonant Auger features in the experimental spectrum, we performed {\it ab initio} calculations on microsolvated clusters of \ki~and \cli. Our calculations show that the energy ordering of the 3d and 4p virtual orbitals of \cli~is inverted compared to \ki, and also that the energy splitting between the bright 1s$^{-1}$4p and dark 1s$^{-1}$3d core excited states is larger in the chloride case. Thus, the energetic proximity of the 3d and 4p orbitals in the bare \ki~ion results in the dipole forbidden 1s$^{-1}$3d state acquiring intensity in a solution as a result of mixing with the dipole allowed 1s$^{-1}$4p excitation. The spectator Auger decay of this state produces an additional dispersive feature which is manifested as a separate peak in the Auger electron spectrum at high kinetic energies. In the case of \cli~the 1s$^{-1}$4p and 1s$^{-1}$3d core excited states do not interact, and therefore, only fingerprints of the population and Auger decay of the dipole allowed 1s$^{-1}$4p excitation are observed in the spectrum. Moreover, using the core-hole clock method we estimated the time of delocalization of the core excited electron at the pre-edge region of \cli$_{\text{aq}}$. Our results show that in this case, the resonant Auger decay and the delocalization of the excited electron occur on a comparable timescale. In the case of \ki$_{\text{aq}}$, an accurate estimate is difficult, however, one can expect a much less efficient delocalization of the core excited electron.

Our work shows that the combination of x-ray absorption and resonant Auger spectroscopy is a sensitive probe of the electronic structure of solvated ions. The reported results are an important first step in the study of the chains of relaxation steps triggered by photoabsorption in the tender x-ray regime. The Auger processes considered here are inevitably followed by multiple intra- and interatomic electronic decays, such as interatomic Coulombic decay (ICD) and electron-transfer mediated decay (ETMD). As a result of the latter processes, genotoxic free radicals and slow electrons are formed in the vicinity of the metal center. The magnitude of the damage inflicted upon the environment and the energies of the emitted electrons depend on the initial Auger step, and can therefore be controlled by tuning the energy of the radiation. Consequently, the results of this work can have implications in understanding radiation chemistry and radiation damage in biologically relevant systems.