### 2018-05-16 jacs_version/abstract.tex

X-ray absorption and Auger electron spectroscopies are powerful tools to probe the electronic structure and immediate surroundings of ions in solution. In this work we use a combination of these methods to study the electronic structure and decay in aqueous KCl at the K-edges of \ki~and \cli. Although the two ions are isoelectronic, their Auger electron spectra as a function of the photon energy exhibit notably different features. To explain these differences, we carried out {\it ab initio} calculations of both the core excited states and the final Auger states of bare \ki, \cli~and their microsolvated clusters. Our calculations show that the energetic order of the 3d and 4p orbitals is inverted in \ki~with respect to \cli. The reverse orbital order in the two ions is reflected in both the ordering of the core excited states and in the final states populated in the resonant Auger decay. The energetic proximity of the 3d and 4p virtual orbitals in the bare \ki~ion leads to their mixing in the presence of the solvent, and to the population of the dipole forbidden 1s$^{-1}$3d state upon K-shell excitation in aqueous solution. The resonant Auger decay of this state results in a separate feature in the Auger electron spectrum of \ki~which is absent in the spectrum of \cli. The results of this work represent a pioneering study of the decay processes initiated by photoabsorption in the tender x-ray regime close to threshold in liquids and are thus of importance in unveiling the mechanisms of radiation damage in biologically relevant systems.
%Denis: line 2, I replaced spectroscopies by methods